### 2023-02-24 code/processing_code/processingfile_v1.qmd
working on cleaning food variable+move EDA part to analysis code

---
title: "Data Pre-Processing/Cleaning Script"
author: "Andreas Handel"
date: "2023-01-03"
output: html_document
---
# Load Libraries

```{r}
# General Pre-Processing
library(tidyverse)
library(dplyr)
library(stringr)

# Data Structure Visualization
library(skimr)

# Setting Paths
library(here)

# Information About NAs
library(dlookr)
```

# Loading Data

```{r}
# Path to Data
data_location <- here("data","raw_data","foodborne_outbreaks.csv")
training_data <- read_csv(data_location)

data_location <- here(
  "data","raw_data","NationalOutbreakPublicDataTool.xlsx")
analysis_data <- readxl::read_excel(data_location)
```

# Check Data

There are several ways we can look at the datasets' structures.

## Training Data
```{r}
str(training_data)
summary(training_data)

# Closer Look at Training Data
head(training_data)
skim(training_data)

# Plotting NAs of Training Data
plot_na_pareto(training_data)
```

It appears that the training data contains the following variables:
General
-   Year, Month (Time)
-   State (Location)
-   Food, Ingredient (Vehicle of Transmission)
-   Species
-   Status
Health Outcomes
-   Illnesses
-   Hospitalizations
-   Fatalities

## Analysis Data
```{r}
str(analysis_data)
summary(analysis_data)

# Closer Look at Training Data
head(analysis_data)
skim(analysis_data)

# Plotting NAs of Training Data
plot_na_pareto(analysis_data)
```

It appears that the training data contains the following variables:
General
-   Year, Month (Time)
-   State (Location)
-   Primary Mode (Vehicle of Transmission)
-   Food Vehicle
-   Food Contaminated Ingredient
-   Water Exposure
-   Water Type
-   Water Status
-   Animal Type
-   Animal Type Specify
-   Etiology
-   Serotype or Genotype
-   Etiology Status
-   Setting
-   IFSAC Category
Health Outcomes
-   Illnesses
-   Hospitalizations
-   Info on Hospitalizations
-   Deaths
-   Info on Deaths

Because the analysis dataset contains more than just foodborne outbreaks, we want to specify it to just foodborne illnesses.

```{r}
# Plot of NAs After Limiting to Analysis Data to Food
plot_na_pareto(analysis_data %>% filter(`Primary Mode` == "Food"))
```

## Checking Etiologies Status
```{r}
# Training Data
unique(training_data$Status)

# Analysis Data
unique(analysis_data$`Etiology Status`)
```

Given that "Status" and "Etiology Status" refer to laboratory confirmations, we will filter our case definitions to be only laboratory-confirmed pathogens.

# Cleaning

Based on the data inspections completed above, we want to remove "Ingredient" and "Serotype/Genotype" from the training dataset (> 80% data missing), and we want to remove all variables unrelated to foodborne illness for the analysis dataset (ex: animal type, water type, etc.) in addition to columns with many NAs.

## Training Data
```{r}
training_data_fin <- training_data %>%
  select(!c(Ingredient, `Serotype/Genotype`)) %>%
  filter(grepl("Confirmed", Status))
```
#Data manipulation
##checking and cleaning the `Species` column
```{r}
unique(training_data_fin$Species)
#Separate the rows for coninfections
training_data_fin=training_data_fin%>%
  separate_rows(Species,sep=";")
#Create lists for pathogen types
Pathogen_Bacteria <- c("Bacillus", "Brucella", "Campylobacter", "Clostridium", "Staphylococcus", "Escherichia", "Listeria", "Salmonella", "Vibrio", "Cholera", "Enterococcus", "Bacterium", "Shigella", "Yersinia","Streptococcus")
Pathogen_Parasite <- c("Cryptosporidium", "Cyclospora", "Entamoeba", "Giardia", "Parasite", "Toxoplasma", "Trichinella")
Pathogen_Virus <- c("Astrovirus", "Norovirus", "Rotavirus","Hepatitis","Virus")
Pathogen_Poison <- c("Poison", "Toxin", "Metal","Pesticide","Ciguatoxin","Scombroidtoxin")
#simplify "Escherichia coli, Shiga toxin-producing" to "Escherichia coli"
#Create two new columns: Pathogen_type & Simplified_Etiology
training_data_fin=training_data_fin%>%
  mutate(Species=str_replace(Species,"Escherichia coli, Shiga toxin-producing","Escherichia"))%>%
  mutate(Pathogen_Type = case_when(
    grepl(paste(Pathogen_Poison, 
                collapse = "|"), Species,ignore.case = TRUE) ~ "Poison",
    grepl(paste(Pathogen_Bacteria, 
                collapse = "|"), Species,ignore.case = TRUE) ~ "Bacteria",
    grepl(paste(Pathogen_Parasite, 
                collapse = "|"), Species,ignore.case = TRUE) ~ "Parasite",
    grepl(paste(Pathogen_Virus, 
                collapse = "|"), Species,ignore.case = TRUE) ~ "Virus"
    )) %>%
  mutate(Simplified_Etiology = case_when(
    Pathogen_Type == "Bacteria" ~ str_extract(Species, 
      "Bacillus|Brucella|Campylobacter|Clostridium|Staphylococcus|Escherichia|Listeria|Salmonella|Vibrio|Cholera|Enterococcus|Bacterium|Shigella|Yersinia|Streptococcus"),
    Pathogen_Type == "Parasite" ~ str_extract(Species, "Cryptosporidium|Cyclospora|Entamoeba|Giardia|Parasite|Toxoplasma|Trichinella"),
    Pathogen_Type == "Virus" ~ str_extract(Species, "Astrovirus|Norovirus|Rotavirus|Hepatitis"),
    Pathogen_Type == "Poison" ~ str_extract(Species, "Poison|Toxin|Heavy|pesticide|Ciguatoxin|Scombroid")
    )) %>%
      mutate(Simplified_Etiology = case_when(
        Simplified_Etiology == "Escherichia" ~ "E. coli",
        Simplified_Etiology == "Heavy" ~ "Heavy Metals",
        TRUE ~ Simplified_Etiology))%>%
  view()
```
# Column for `Location` 
```{r}
#checking the names of location and separate the collapsed names
unique(training_data_fin$Location)
training_data_fin=training_data_fin%>%
 separate_rows(Location,sep=";")%>%
  view()

```
# Location variable looks clean
### Add Month as Numeric Column
```{r}
training_data_fin <- training_data_fin %>%
  mutate(Numeric_Month = case_when(
    Month == "January" ~ 1,
    Month == "February" ~ 2,
    Month == "March" ~ 3,
    Month == "April" ~ 4,
    Month == "May" ~ 5,
    Month == "June" ~ 6,
    Month == "July" ~ 7,
    Month == "August" ~ 8,
    Month == "September" ~ 9,
    Month == "October" ~ 10,
    Month == "November" ~ 11,
    Month == "December" ~ 12)) %>%
  relocate(Numeric_Month, .after = Month)
```
##Since there are over 3000+ different categories for food, we want to organize them by adding a IFSAC category column
```{r}
unique(analysis_data_fin$`IFSAC Category`)
unique(training_data_fin$Food)
Desserts=c("Cookie","Cream","Canoli","Cake","Pudding")
Beef=c("Beef","Steak")
Poultry=c("Chicken","Turkey")
Pork=c("Pork","Sausage","Hot Dog","Deli","Spam","Strawberries","Blueberries","Pig","Ribs")
Dairy=c("Milk","Cheese","Parsley")
Crusteaceans=c("Lobster","Crab","Shrimp")
Fish=c("Fish","Tuna","Salmon")
Mollusks=c("Clams","Oysters")
Fruits=c("Tomato","Banana","Melon","Grapes","Mango")
`Grains-Beans`=c("Rice","Peas","Beans","wheat")
`Vegetable-Row-Crops`=c("Lettuce","artichokes","broccoli","asparagus","celeries","spinach")
`Root-underground-vegetables`=c("Roots","carrots","Beets","Onion","Tubers","yams","Garlic","Potatos","taro")
Sprouts=c("alfalfa","mung")
Salad=c("Salad")
Other=c("Ice","Alcohol","Eggnog")
Game=c("Bear")
Herbs=c("basil","cilantro")
Fungi=c("Mushroom","portabellas")
`Nuts-Seeds`=c("peanuts","almonds","Seasames")
Eggs=c("Egg")
`Seeded-Vegetables`=c("Peppers","Squashes","Cucumbers","Okras","Corn")
`other meat`=c("Lamb")

training_data_fin <- training_data_fin %>%
  mutate(`IFSAC Category` = case_when(
    grepl("Salad",Food, ignore.case = TRUE) ~ "Salad",
    grepl(paste(Desserts, 
                collapse = "|"), Food,ignore.case = TRUE) ~ "Dessert",
    grepl(paste(Poultry,
                collapse = "|"), Food, ignore.case = TRUE) ~ "Poultry",
    grepl(paste(Beef,
                collapse = "|"), Food, ignore.case = TRUE) ~ "Beef",
    grepl(paste(Pork,
                collapse = "|"), Food, ignore.case = TRUE) ~ "Pork",
    grepl(paste(Dairy,
                collapse = "|"), Food, ignore.case = TRUE) ~ "Dairy",
    grepl(paste(Crusteaceans,
                collapse = "|"), Food, ignore.case = TRUE) ~ "Crusteaceans",
    grepl(paste(Fish,
                collapse = "|"), Food, ignore.case = TRUE) ~ "Fish",
    grepl(paste(Mollusks,
                collapse = "|"), Food, ignore.case = TRUE) ~ "Mollusks",
    grepl(paste(Fruits,
                collapse = "|"), Food, ignore.case = TRUE) ~ "Fruits",
    grepl(paste(`Grains-Beans`,
                collapse = "|"), Food, ignore.case = TRUE) ~ "Grains-Beans",
    grepl(paste(`Vegetable-Row-Crops`,
                collapse = "|"), Food, ignore.case = TRUE) ~ "Vegetable-Row-Crops",
    grepl(paste(`Root-underground-vegetables`,
                collapse = "|"), Food, ignore.case = TRUE) ~ "Root-underground-vegetables",
    grepl(paste(Sprouts,
                collapse = "|"), Food, ignore.case = TRUE) ~ "Sprouts",
    grepl(paste(Other,
                collapse = "|"), Food, ignore.case = TRUE) ~ "Other",
    grepl(paste(Herbs,
                collapse = "|"), Food, ignore.case = TRUE) ~ "Herbs",
    grepl(paste(Fungi,
                collapse = "|"), Food, ignore.case = TRUE) ~ "Fungi",
    grepl(paste(`Nuts-Seeds`,
                collapse = "|"), Food, ignore.case = TRUE) ~ "Nuts-Seeds",
    grepl(paste(`Seeded-Vegetables`,
                collapse = "|"), Food, ignore.case = TRUE) ~ "Seeded-Vegetables",
    grepl("eggs",Food,ignore.case = TRUE)~"Eggs"

   ))%>%
  #mutate(`IFSAC Category`=case_when(Food==str_detect(Food, ";")~"Multiple"))%>%
  view()
```

## Analysis Data
```{r}
analysis_data_fin <- analysis_data %>%
  select(!c(`Animal Type`, `Animal Type Specify`, 
            `Water Exposure`, `Water Type`, `Water Status`, 
            `Food Contaminated Ingredient`)) %>%
  filter(grepl("Confirmed", `Etiology Status`)) %>%
  filter(`Primary Mode` == "Food")
```

## Data Manipulation
### Separate Rows to Address Coinfections
```{r}
analysis_data_fin <- analysis_data_fin %>%
  separate_rows(Etiology, sep = "; ")
```

### Modifying IFSAC Category to Include Salads
```{r}
Desserts <- c("cookies", "cake", "chocolate", "brownie")

analysis_data_fin <- analysis_data_fin %>%
  mutate(`IFSAC Category` = case_when(
    grepl("Salad", `Food Vehicle`, ignore.case = TRUE) ~ "Salad",
    grepl(paste(Desserts, 
                collapse = "|"), `Food Vehicle`) ~ "Dessert",
    grepl("Chicken", `Food Vehicle`, ignore.case = TRUE) ~ "Poultry",
    grepl("Turkey", `Food Vehicle`, ignore.case = TRUE) ~ "Poultry",
    grepl("Poultry", `Food Vehicle`, ignore.case = TRUE) ~ "Poultry",
    TRUE ~ `IFSAC Category`))
```

### Column for Type of Pathogen
```{r}

analysis_data_fin <- analysis_data_fin %>%
  mutate(Pathogen_Type = case_when(
    grepl(paste(Pathogen_Poison, 
                collapse = "|"), Etiology) ~ "Poison",
    grepl(paste(Pathogen_Bacteria, 
                collapse = "|"), Etiology) ~ "Bacteria",
    grepl(paste(Pathogen_Parasite, 
                collapse = "|"), Etiology) ~ "Parasite",
    grepl(paste(Pathogen_Virus, 
                collapse = "|"), Etiology) ~ "Virus"
    )) %>%
  mutate(Simplified_Etiology = case_when(
    Pathogen_Type == "Bacteria" ~ str_extract(Etiology, 
      "Bacillus|Brucella|Campylobacter|Clostridium|Staphylococcus|Escherichia|Listeria|Salmonella|Vibrio|Cholera|Enterococcus|Bacterium|Shigella|Yersinia"),
    Pathogen_Type == "Parasite" ~ str_extract(Etiology, "Cryptosporidium|Cyclospora|Entamoeba|Giardia|Parasite|Toxoplasma|Trichinella"),
    Pathogen_Type == "Virus" ~ str_extract(Etiology, "Astrovirus|Norovirus|Rotavirus"),
    Pathogen_Type == "Poison" ~ str_extract(Etiology, "Poison|Toxin|Heavy")
    )) %>%
      mutate(Simplified_Etiology = case_when(
        Simplified_Etiology == "Escherichia" ~ "E. coli",
        Simplified_Etiology == "Heavy" ~ "Heavy Metals",
        TRUE ~ Simplified_Etiology))
```
#we want to drop the NA rows
```{r}
training_data_fin%>%
  drop_na()
analysis_data_fin%>%
  drop_na()
```

## Check Data Structure Again
```{r}
skim(training_data_fin)
skim(analysis_data_fin)
```

All done, data is clean now.

Let's assign at the end to some final variable, this makes it easier to add further cleaning steps above.

```{r}
processeddata1<- training_data_fin
processeddata2<-analysis_data_fin
```

# Save data

Finally, we save the clean data as RDS files.

```{r}
save_data_location1 <- here::here("data","processed_data","processeddata1.rds")
saveRDS(processeddata1, file = save_data_location1)

save_data_location2=here::here("data","processed_data","processeddata2.rds")
saveRDS(processeddata2, file = save_data_location2)

```

# Notes

Removing anyone who had "faulty" or missing data is one approach. It's often not the best. based on your question and your analysis approach, you might want to do cleaning differently (e.g. keep individuals with some missing information).