### 2023-03-14 code/analysis_code/exploratory_analysis.qmd
update processing and exploratory analysis files

---
title: "An example exploratory analysis script"
author: "Andreas Handel"
date: "10/28/2021"
output: html_document
---

# Load Libraries
```{r}
# General Data Handling
library(tidyverse)

# For Data Loading and Saving
library(here)
```

# Load Data

## Training Data
```{r}
# Path to Data
data_location <- here(
  "data","processed_data","processed_training_data.rds")
# Import Data
training_data_final <- readRDS(data_location)
```

## Analysis Data
```{r}
# Path to Data
data_location <- here(
  "data","processed_data","processed_testing_data.rds")
# Import Data
analysis_data_final <- readRDS(data_location)
```

# Data Exploration
## Training Data

The outcome of interest we will explore for the training data is hospitalizations.

```{r}
# Time vs Hospitalizations
ggplot(training_data_final %>% 
         group_by(Year, Pathogen_Type, Hospitalizations) %>% 
         count() %>% 
         mutate(true_hospitalization = n * Hospitalizations), aes(x = Year, y = true_hospitalization, group = Pathogen_Type, color = Pathogen_Type)) + geom_point() + 
  geom_jitter() + 
  labs(x = "Year", y = "Hospitalizations", title = "Scatter Plot of Hospitalizations by Year and Pathogen Type")

## Save Graph
image_path <- here("results")
ggsave("time_v_hospitalization_pathogen_type.png", path = image_path, width = 3000, height = 2000, units = "px")

# Pathogen Type + Etiology vs Hospitalization
ggplot(training_data_final %>% 
         group_by(Pathogen_Type, Simplified_Etiology, Hospitalizations) %>% count() %>% 
         mutate(true_hospitalization = n * Hospitalizations), aes(x = Pathogen_Type, y = true_hospitalization, group = Simplified_Etiology, fill = Simplified_Etiology)) +
  geom_col() + 
  labs(x = "Pathogen Type", y = "Hospitalizations", title = "Stacked Bar Chart of Hospitalizations by Pathogen Type and Etiology")

## Save Graph
ggsave("pathogen_type_etiology_v_hospitalizations.png", path = image_path, width = 3000, height = 2000, units = "px")

## Closer Look at Bacterial Etiologies for Outbreaks
ggplot(training_data_final %>% 
         select(Pathogen_Type, Simplified_Etiology, Hospitalizations, `IFSAC Category`) %>% filter(Pathogen_Type == "Bacteria"), aes(x = Simplified_Etiology,y=`Hospitalizations`))+ 
  geom_col() + 
  labs(x = "Bacteria", y = "Hospitalizations", title = "Bar Chart of Bacteria Frequency")+
  coord_flip()

ggsave("bacterial_etiology_v_hospitalization_frequency.png", path = image_path, width = 3000, height = 2000, units = "px")

# States vs Hospitalization

state_v_hospitalization_training <- training_data_final %>% 
  group_by(State) %>% 
  mutate(all_hospitalizations = sum(Hospitalizations)) %>%     count(all_hospitalizations) %>% 
  select(State, all_hospitalizations) %>% arrange(desc(all_hospitalizations))

## Save Table
table_path <- here("results", "state_v_hospitalization_training.rds")
saveRDS(state_v_hospitalization_training, file = table_path)

rmarkdown::paged_table(state_v_hospitalization_training)

## Project total hospitalizations in states map
training_hosp=training_data_final%>%
  mutate(state=tolower(State))%>%
  group_by(state)%>%
  mutate(all_hosp = sum(Hospitalizations)) %>%  count(all_hosp) %>%
  select(state, all_hosp)
training_hosp

state_map=training_hosp%>%
  distinct(state)%>%
  inner_join(.,map_data("state"),by=c("state"="region"))
state_map

map_final=right_join(training_hosp,state_map,by=c("state"="state"))

map_final%>%
  ggplot(aes(x=long,y=lat,fill=all_hosp,group=group))+
  geom_polygon(color="black",size=0.1)+
  geom_polygon(data=state_map,aes(x=long,y=lat,group=group),fill=NA,color="black",size=0.3)+
  scale_fill_gradient(name="Number of \n Hospitalizations",
                      low="#FFFFFF",high="#FF0000",
                       limits=c(0,NA))+
  labs(title="Cases of Hospitalizations Caused by Foodborne Illnesses\n in Each State During 1998-2015")+
  theme_void()+
  coord_quickmap()

ggsave("State_hospitalization_map.png",path=image_path,width=3000,height=2000,units="px")

# Food Type and Hospitalizations
ggplot(training_data_final %>% 
         filter(`IFSAC Category` != "Multiple") %>% 
         filter(`IFSAC Category` != "Other"), 
       aes(x = `IFSAC Category`, y = Hospitalizations)) + 
  geom_col() + 
  labs(x = "Food Type", y = "Hospitalizations", title = "Hospitalizations by Food Type, Excluding 'Multiple' and 'Other'")+
  coord_flip()

ggsave("Food_type_v_hospitalization.png", path = image_path, width = 3000, height = 2000, units = "px")

# Locations and Hospitalizations
Location_v_Hospitalization <- ggplot(training_data_final, aes(x = Location, y = Hospitalizations)) +
  geom_col() + 
  labs(x = "Locations", y = "Hospitalizations", title = "Hospitalizations by Food Type, Excluding 'Multiple' and 'Other'")+
  coord_flip()

Location_v_Hospitalization

ggsave("Location_v_hospitalization.png", path = image_path, width = 3000, height = 2000, units = "px")

```
All outbreaks with a high hospitalization count (> 200 cases) are caused by bacteria. It also appears that most outbreaks caused by bacteria are due to Salmonella. California and Minnesota are the two states with the highest total hospitalization count due to foodborne illness from 1998 to 2015. Resturants and Private home are the most frequent locations with the highest total hospitalization count.

## Testing Data
```{r}
# Illnesses Caused by Pathogen Type Over Time
ggplot(testing_data_final %>% 
         group_by(Year, Pathogen_Type, Illnesses) %>% 
         count() %>% 
         mutate(true_illness_count = n * Illnesses), 
       aes(x = Year, y = true_illness_count, group = Pathogen_Type, color = Pathogen_Type)) + 
  geom_point() + 
  geom_jitter() + 
  labs(x = "Year", y = "Cases", title = "Scatter Plot of Illness by Year and Pathogen Type")

## Save Graph
ggsave("time_v_illness_pathogen_type.png", path = image_path, width = 3000, height = 2000, units = "px")

# Pathogen Type + Etiology vs Illnesses Caused
ggplot(testing_data_final %>% 
         group_by(Pathogen_Type, Simplified_Etiology, Illnesses) %>% 
         count() %>% 
         mutate(true_illness_count = n * Illnesses), 
       aes(x = Pathogen_Type, y = Illnesses, group = Simplified_Etiology, fill = Simplified_Etiology)) + 
  geom_col() + 
  labs(x = "Pathogen Type", y = "Cases", title = "Stacked Bar Chart of Illnesses by Pathogen Type and Etiology")

## Save Graph
ggsave("pathogen_type_etiology_v_illness.png", path = image_path, width = 3000, height = 2000, units = "px")

## Closer Look at Bacterial Etiologies
ggplot(testing_data_final %>% group_by(Pathogen_Type, Simplified_Etiology, Illnesses) %>% filter(Pathogen_Type == "Bacteria"), aes(x = Simplified_Etiology)) + geom_bar() + labs(x = "Bacteria", y = "Cases", title = "Bar Chart of Bacteria Frequency")+
  coord_flip()

# Save Graph
ggsave("bacterial_etiology_v_outbreak_frequency.png", path = image_path, width = 3000, height = 2000, units = "px")

# States vs Illnesses

state_v_illness_testing <- testing_data_final %>% group_by(State) %>% mutate(all_illnesses = sum(Illnesses)) %>% count(all_illnesses) %>% select(State, all_illnesses) %>% arrange(desc(all_illnesses))

saveRDS(state_v_illness_testing, file = table_path)

rmarkdown::paged_table(state_v_illness_analysis)

# Food Type and Illnesses
ggplot(testing_data_final %>% filter(`IFSAC Category` != "Multiple") %>% filter(`IFSAC Category` != "Other"), aes(x = `IFSAC Category`, y = Illnesses)) + geom_col() + labs(x = "Food Type", y = "Illnesses", title = "Illnesses by Food Type, Excluding 'Multiple' and 'Other'")+
  coord_flip()

# Save Graph
ggsave("food_type_v_illness_frequency.png", path = image_path, width = 3000, height = 2000, units = "px")

# Locations and Illnesses
ggplot(testing_data_final, aes(x = Location, y = Illnesses)) +
  geom_col() + 
  labs(x = "Locations", y = "Illnesses", title = "Illnesses by Locations")+
  coord_flip()

# Save Graph
ggsave("Location_v_Illness.png", path = image_path, width = 3000, height = 2000, units = "px")

```
The outbreaks with a high number of cases (> 500) are caused by bacteria. The stacked barplot indicates that the top two etiologic agents that contribute to fooborne-illnesses are Salmonella and Norovirus, which belong to the pathogen type of Bacteria and Virus respectively. In bacterial pathogen, Salmonella genus is the most dominant factor that is associated with foodborne outbreaks, followed by E. coli species and Vibrio genus. Geographically speaking, Arkansas and Minnesota are the two states with the highest accumulated foodborne illnesses from 1998 to 2015. Salad, Poultry, and Fruits are the three most common types of food vectors that are involved in transmitting foodborne pathogens. Resturants and Private home are the most frequent locations that foodborne outbreaks occur.
Overall, the relationships between predictors and outcomes of training dataset share similar patterns with those of analysis dataset.