### 2023-03-16 code/analysis_code/statistical_analysis.qmd

# Load libraries
```{r}
library(here)
library(tidyverse)
library(tidymodels)
library(poissonreg)
```
# Load data
```{r}
# Path to Data
data_location <- here(
  "data","processed_data","processed_training_data.rds")
# Import Data
training_data_final <- readRDS(data_location)
glimpse(training_data_final)
```
## First health outcome of interest: Illnesses
### Predictors: time (year), pathogen type, etiology, location, state, and food type (`IFSAC Category`)
```{r}
# Checking the distribution of Illnesses
training_data_final%>%
  ggplot(aes(Illnesses))+
  geom_histogram()+
  scale_x_continuous(trans=scales::pseudo_log_trans(base=10))
image_path <- here("results")
ggsave("histogram_Illnesses.png", path = image_path, width = 3000, height = 2000, units = "px")
# Creating a tibble only includes our predictors of interest
illness_data=training_data_final%>%
  select(c(1,4,5,7,10,13,14))%>%
  mutate_if(is.character, as.factor)
glimpse(illness_data)
```
### Modeling `Illness` using linear regression and Poisson regression
#### Since the outcome is counts and the event ocurrs independently, it follows a Poisson distribution
```{r}
# Model specification and fitting for linear regression model
lm_spec=linear_reg()%>%
  set_engine(engine="lm")
# Modeling with main predictor `Simplified_Etiology`
lm_fit1=lm_spec%>%
  fit(Illnesses~Simplified_Etiology, data=illness_data)
# Modeling with all predictors
lm_fit2=lm_spec%>%
  fit(Illnesses~., data=illness_data)
# Model specification and fitting for Poisson regression model
# create a workflow for poisson regression model, we will use the default glm engine
# Modeling with main predictor `Simplified_Etiology`
poisson_wf1=workflow(Illnesses~Simplified_Etiology, poisson_reg())
# Modeling with all predictors
poisson_wf2=workflow(Illnesses~., poisson_reg())
# Fitting models
poisson_fit1=fit(poisson_wf1,data=illness_data)
poisson_fit2=fit(poisson_wf2,data=illness_data)
```

### Model Evaluation
```{r}
# Checking parameters for linear regression model
tidy(lm_fit1) %>%
  arrange(-estimate)
tidy(lm_fit2)%>%
  arrange(-estimate)
# Checking parameters for Poisson regression model
tidy(poisson_fit1)%>%
 arrange(-estimate)
tidy(poisson_fit2)%>%
 arrange(-estimate)
# Model evaluation
Result=lm_fit1%>%
 predict(new_data=illness_data)%>%
 mutate (.pred=round(.pred),true_value=illness_data$Illnesses, model= "lm_single")%>%
bind_rows(lm_fit1%>%
            predict(new_data=illness_data)%>%
            mutate (.pred=round(.pred),true_value=illness_data$Illnesses, model= "lm_multiple")%>%
            bind_rows(poisson_fit1%>%
                        predict(new_data= illness_data)%>% 
                        mutate (true_value=illness_data$Illnesses, model= "glm_single")%>%
                        bind_rows(poisson_fit2%>%
                                    predict(new_data= illness_data)%>% 
                                    mutate (true_value=illness_data$Illnesses, model= "glm_multiple")
                        )
            )
)
Result%>%
  group_by(model)%>%
  rmse(truth=true_value,estimate=.pred)

```
### According to the result, the model built using Poisson regression with multiple predictors seems to have the smallest root mean squared error.
## Second health outcome of interest: Hospitilization
### Predictors: time (year), pathogen type, etiology, location, state, and food type (`IFSAC Category`)
```{r}
# Checking the distribution of Hospitalization
training_data_final%>%
  ggplot(aes(Hospitalizations))+
  geom_histogram()+
  scale_x_continuous(trans=scales::pseudo_log_trans(base=10))
# Creating a tibble only includes our predictors of interest
hosp_data=training_data_final%>%
  select(c(1,4,5,7,11,13,14))%>%
  mutate_if(is.character, as.factor)
glimpse(hosp_data)
ggsave("histogram_Hospitalizations.png",path=image_path)
```
### Modeling `Hospitalizations` using linear regression and Poisson regression
```{r}
# Model specification and fitting for linear regression model
lm_spec=linear_reg()%>%
  set_engine(engine="lm")
# Modeling with main predictor `Simplified_Etiology`
lm_fit3=lm_spec%>%
  fit(Hospitalizations~Simplified_Etiology, data=hosp_data)
# Modeling with all predictors
lm_fit4=lm_spec%>%
  fit(Hospitalizations~., data=hosp_data)
# Model specification and fitting for Poisson regression model
# create a workflow for poisson regression model, we will use the default glm engine
# Modeling with main predictor `Simplified_Etiology`
poisson_wf3=workflow(Hospitalizations~Simplified_Etiology, poisson_reg())
# Modeling with all predictors
poisson_wf4=workflow(Hospitalizations~., poisson_reg())
# Fitting models
poisson_fit3=fit(poisson_wf3,data=hosp_data)
poisson_fit4=fit(poisson_wf4,data=hosp_data)
```

### Model evaluation
```{r}
Result=lm_fit3%>%
 predict(new_data=hosp_data)%>%
 mutate (.pred=round(.pred),true_value=hosp_data$Hospitalizations, model= "lm_single")%>%
bind_rows(lm_fit4%>%
            predict(new_data=hosp_data)%>%
            mutate (.pred=round(.pred),true_value=hosp_data$Hospitalizations, model= "lm_multiple")%>%
            bind_rows(poisson_fit1%>%
                        predict(new_data= hosp_data)%>% 
                        mutate (true_value=hosp_data$Hospitalizations, model= "glm_single")%>%
                        bind_rows(poisson_fit2%>%
                                    predict(new_data= hosp_data)%>% 
                                    mutate (true_value=hosp_data$Hospitalizations, model= "glm_multiple")
                        )
            )
)
Result%>%
  group_by(model)%>%
  rmse(truth=true_value,estimate=.pred)

```