### 2023-03-16 code/analysis_code/statistical_analysis.qmd
update

# Load libraries
```{r}
# Pathing
library(here)

# Data Handling and Modeling
library(tidyverse)
library(tidymodels)
library(poissonreg)
```
# Load Data
```{r}
# Path to Data
data_location <- here(
  "data","processed_data","processed_training_data.rds")
# Import Data
training_data_final <- readRDS(data_location)
glimpse(training_data_final)
```
# First Health Outcome of Interest: Illnesses

Predictors
-   Time (Year)
-   Pathogen Type
-   Etiology
-   Location
-   State
-   Food Type (IFSAC Category)

```{r}
# Checking Illness Distribution
training_data_final %>%
  ggplot(aes(Illnesses))+
  geom_histogram()+
  scale_x_continuous(trans=scales::pseudo_log_trans(base=10))
image_path <- here("results")
ggsave("histogram_Illnesses.png", path = image_path, width = 3000, height = 2000, units = "px")

# Tibble with Predictors of Interest
illness_data=training_data_final %>%
  select(c(1, 4, 5, 7, 10, 13, 14)) %>%
  mutate_if(is.character, as.factor)
glimpse(illness_data)
```

## Modeling Illness with Linear Regression and Poisson Regression

Since the outcome is counts/frequency and the event occurs independently, it follows a Poisson distribution. To create a workflow for poisson regression model, the default *glm* engine will be used.

```{r}
# Linear Regression
lm_spec = linear_reg() %>%
  set_engine(engine = "lm")
## Modeling with Main Predictor `Simplified_Etiology`
lm_fit1 = lm_spec %>%
  fit(Illnesses ~ Simplified_Etiology, data = illness_data)
## Modeling with All Predictors
lm_fit2 = lm_spec %>%
  fit(Illnesses ~ ., data = illness_data)

# Poisson Regression
## Modeling with Main Predictor `Simplified_Etiology`
poisson_wf1 = workflow(Illnesses ~ Simplified_Etiology, poisson_reg())
## Modeling with All Predictors
poisson_wf2 = workflow(Illnesses ~ ., poisson_reg())
## Model Fitting
poisson_fit1 = fit(poisson_wf1, data=illness_data)
poisson_fit2 = fit(poisson_wf2, data=illness_data)
```

### Model Evaluation
```{r}
# Checking Parameters for Linear Regression Model
tidy(lm_fit1) %>%
  arrange(-estimate)
tidy(lm_fit2) %>%
  arrange(-estimate)

# Checking Parameters for Poisson Regression Model
tidy(poisson_fit1) %>%
 arrange(-estimate)
tidy(poisson_fit2) %>%
 arrange(-estimate)

# Model Evaluation
Result = lm_fit1 %>% 
  predict(new_data = illness_data) %>% 
  mutate(.pred = round(.pred), 
        true_value = illness_data$Illnesses, model= "lm_single") %>% 
  bind_rows(lm_fit1 %>% 
              predict(new_data=illness_data) %>% 
              mutate(.pred=round(.pred), 
                     true_value = illness_data$Illnesses, model= "lm_multiple") %>% 
              bind_rows(poisson_fit1 %>% predict(new_data= illness_data) %>%
                          mutate(true_value=illness_data$Illnesses, 
                                 model= "glm_single") %>% 
                          bind_rows(poisson_fit2 %>% 
                                      predict(new_data= illness_data)%>% 
                                      mutate(true_value=illness_data$Illnesses, 
                                             model= "glm_multiple")))
)

Result <- Result %>%
  group_by(model) %>%
  rmse(truth=true_value, estimate = .pred) %>%
  rename(illness_model = model)

saveRDS(Result, file = here("results", "illness_model_evaluation.rds"))

```
According to the result, the model built using Poisson regression with multiple predictors seems to have the smallest root mean squared error.

# Second Health Outcome of Interest: Hospitilization

Predictors
-   Time (Year)
-   Pathogen Type
-   Etiology
-   Location
-   State
-   Food Type (IFSAC Category)

```{r}
# Checking the Distribution of Hospitalization
training_data_final %>%
  ggplot(aes(Hospitalizations)) +
  geom_histogram() +
  scale_x_continuous(trans=scales::pseudo_log_trans(base = 10))
# Tibble for Variables of Interest
hosp_data = training_data_final %>%
  select(c(1, 4, 5, 7, 11, 13, 14))%>%
  mutate_if(is.character, as.factor)
glimpse(hosp_data)
ggsave("histogram_Hospitalizations.png", path = image_path)
```

## Modeling `Hospitalizations` using Linear Regression and Poisson Regression

To create a workflow for the Poisson regression model, we will use the default *glm* engine again.

```{r}
# Model Specification and Fitting for Linear Regression Model
lm_spec = linear_reg() %>%
  set_engine(engine = "lm")
## Modeling with Main Predictor `Simplified_Etiology`
lm_fit3 = lm_spec %>%
  fit(Hospitalizations ~ Simplified_Etiology, data = hosp_data)
## Modeling with All Predictors
lm_fit4 = lm_spec %>%
  fit(Hospitalizations ~ ., data = hosp_data)

# Model Specification and Fitting for Poisson Regression Model
## Modeling with Main Predictor `Simplified_Etiology`
poisson_wf3 = workflow(Hospitalizations ~ Simplified_Etiology, poisson_reg())
## Modeling with all predictors
poisson_wf4=workflow(Hospitalizations~., poisson_reg())
## Fitting models
poisson_fit3 = fit(poisson_wf3,data = hosp_data)
poisson_fit4 = fit(poisson_wf4,data = hosp_data)
```

### Model Evaluation
```{r}
Result = lm_fit3 %>%
 predict(new_data = hosp_data) %>%
 mutate(.pred = round(.pred), true_value = hosp_data$Hospitalizations,
        model= "lm_single") %>% 
  bind_rows(lm_fit4 %>% 
              predict(new_data=hosp_data) %>% 
              mutate (.pred=round(.pred), 
                      true_value=hosp_data$Hospitalizations, 
                      model= "lm_multiple") %>% 
              bind_rows(poisson_fit1 %>% 
                        predict(new_data = hosp_data) %>%
                        mutate(true_value = hosp_data$Hospitalizations, 
                               model= "glm_single") %>% 
                          bind_rows(poisson_fit2 %>% 
                                      predict(new_data= hosp_data) %>% 
                                      mutate(true_value = hosp_data$Hospitalizations, 
                                             model= "glm_multiple")))
)

Result <- Result %>%
  group_by(model)%>%
  rmse(truth = true_value, estimate = .pred) %>%
  rename(hospital_model = model)

saveRDS(Result, file = here("results", "hospital_model_evaluation.rds"))

```

According to the result, the model built using linear regression with multiple predictors seems to have the smallest root mean squared error.