### 2023-04-07 code/analysis_code/statistical_analysis.qmd

---
title: "Statistical analysis script"
author: "Weifan Wu and Kailin Chen"
date: "03/16/2023"
output: html_document
---

# Load Libraries
```{r}
# Pathing
library(here)

# Data Handling and Modeling
library(tidyverse)
library(tidymodels)
library(poissonreg)
library(knitr)
library(skimr)
library(vip)
## For Decision Trees
library(ranger)
library(caret)
library(rpart)
library(ipred)
library(e1071)
```
# Load Data
```{r}
# Import Training Data
data_location <- here(
  "data","processed_data","processed_training_data.rds")
training_data_final <- readRDS(data_location)
skim(training_data_final)

# Import Test Data
data_location <- here(
  "data","processed_data","processed_testing_data.rds")
test_data_final <- readRDS(data_location)
skim(test_data_final)

```

```{r}
# Set Image Path
images_stats_path <- here("results","images_statistic_analysis")

# Checking General Illness Distribution
training_data_final %>%
  ggplot(aes(Illnesses)) +
  geom_histogram() +
  scale_x_continuous(trans = scales::pseudo_log_trans(base = 10))
## Save Histogram
ggsave("histogram_Illnesses.png", path = images_stats_path, width = 3000, height = 2000, units = "px")

# Checking General Hospitalization Distribution
training_data_final %>%
  ggplot(aes(Hospitalizations)) +
  geom_histogram() +
  scale_x_continuous(trans=scales::pseudo_log_trans(base = 10))
## Save Generated Histogram
ggsave("histogram_Hospitalizations.png", path = images_stats_path)
```

# Create Tibbles for Illness & Hospitalizations
```{r}
process_datasets_to_tibbles <- function(dataset) {
  dataset %>%
    filter(!`IFSAC Category` %in% c("Multiple", "Other"), 
           Location!= "Unknown") %>%
    mutate_if(is.character, as.factor) %>%
    select(!Pathogen_Type)
}

illness_data <- process_datasets_to_tibbles(training_data_final) %>%
  select(!Month) %>%
  rename(Month = Numeric_Month) %>%
  select(!Food) %>% select(!Hospitalizations)

hosp_data <- process_datasets_to_tibbles(training_data_final) %>%
  select(!Month) %>%
  rename(Month = Numeric_Month) %>%
  select(!Food) %>% select(!Illnesses)

skim(illness_data)
skim(hosp_data)

test_data_final <- process_datasets_to_tibbles(test_data_final) %>%
  select(!Food)

skim(test_data_final)
```

The two health outcomes of interest are illnesses and hospitalizations.

Predictors
-   Time (Month, Year)
-   Pathogen Type
-   Etiology
-   Location
-   State
-   Food Type (IFSAC Category)

# Modeling Illness

Since the outcome is counts/frequency and the event occurs independently, it follows a Poisson distribution. To create a workflow for poisson regression model, the default *glm* engine will be used.

```{r}
# Set Seed for Reproducibility
set.seed(123)

# Cross-Validation Setup
illness_fold <- vfold_cv(illness_data, repeats = 2, strata = Illnesses)

# Additional Preprocessing to Deal with Variables with > 50 Categories
preprocessing_recipe <- function(anterior) {
  anterior %>%
    step_dummy(Simplified_Etiology) %>%
    step_dummy(State) %>%
    step_dummy(`IFSAC Category`)
}

## Simple Recipe for Illnesses
illness_simple_recipe <- recipe(Illnesses ~ Simplified_Etiology, 
                                data = illness_data) %>%
  step_dummy(Simplified_Etiology)

## Recipe with All Predictors for Illnesses
illness_complex_recipe <- recipe(Illnesses ~ ., data = illness_data) %>%
  preprocessing_recipe()

### Lump small proportion into "other" using step_other
### Create dummy code for nominal variables
#illness_all_rec=recipe(Illnesses~.,data=illness_data)%>%
#  step_other(State,Location,`IFSAC Category`,Simplified_Etiology,threshold = 0.01)%>%
#  step_dummy(all_nominal())

### Check Processed Dataset
illness_complex_recipe %>% prep() %>% bake(new_data = NULL)

# Model Specification
## Linear Regression
linear_model <- linear_reg() %>%
  set_engine(engine = "lm")

# Create Workflow
## Main Predictor `Simplified_Etiology`
linear_workflow_simple_illness <- workflow() %>%
  add_recipe(illness_simple_recipe) %>%
  add_model(linear_model)

## All Predictors
linear_workflow_complex_illness <- workflow() %>%
  add_recipe(illness_complex_recipe)%>%
  add_model(linear_model)

## Poisson Regression
## Modeling with Main Predictor `Simplified_Etiology`
poisson_workflow_simple_illness <- workflow() %>%
  add_recipe(illness_simple_recipe) %>%
  add_model(poisson_reg())

## Modeling with All Predictors
poisson_workflow_complex_illness <- workflow() %>%
  add_recipe(illness_complex_recipe) %>%
  add_model(poisson_reg())
```

## Linear Regression
### Model Fitting
```{r}
set.seed(234)

## Linear Regression
### Simple Model
lm_simple_resampled_illness <- linear_workflow_simple_illness %>%
  fit_resamples(resamples = illness_fold, control = control_resamples(save_pred = TRUE))

### All Predictors
lm_complex_resampled_illness <- linear_workflow_complex_illness %>%
  fit_resamples(resamples = illness_fold, control = control_resamples(save_pred = TRUE))
```

### Model Evaluation
```{r}
# Check Metrics
## Simple Regression
lm_simple_resampled_illness %>%
  collect_metrics()
## Multivariate Regression
lm_complex_resampled_illness %>%
  collect_metrics()

# Check Parameters for Linear Regression Model
## Simple Regression
linear_model_simple_fit_illness <- linear_workflow_simple_illness %>%
  fit(illness_data) %>%
  tidy() %>%
  arrange(-estimate)
## Multivariate Regression
linear_model_complex_fit_illness <- linear_workflow_complex_illness %>%
  fit(illness_data) %>%
  tidy() %>%
  arrange(-estimate)

# Check Variable Importance
## Simple Regression
linear_workflow_simple_illness %>%
  fit(illness_data) %>%
  extract_fit_engine() %>%
  vip()
## Multivariate Regression
linear_workflow_complex_illness %>%
  fit(illness_data) %>%
  extract_fit_engine() %>%
  vip()

# Plotting Predicted Values vs True Values (Illnesses)
lm_simple_resampled_illness %>%
  collect_predictions() %>%
  # filter(Illnesses < 300) %>%
  ggplot(aes(Illnesses, .pred, color = id2, group = id2)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  geom_abline(intercept = 0,slope = 1) +
  labs(y = "Predicted Illnesses", 
       title = "Predicted Illnesses VS Illnesses via Simple Linear Regression")

ggsave("Predicted Illnesses VS Illnesses by linear regression with main predictor.png", path = images_stats_path, width = 3000, height = 2000, units = "px")

lm_complex_resampled_illness %>%
  collect_predictions() %>%
  # filter(Illnesses < 300) %>%
  ggplot(aes(Illnesses, .pred, color = id2, group = id2)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  geom_abline(intercept = 0, slope = 1) +
  labs(y = "Predicted Illnesses", title = "Predicted Illnesses VS Illnesses via Multivariate Linear Regression")

ggsave("Predicted Illnesses VS Illnesses by linear regression with all predictors.png", path=images_stats_path, width = 3000, height = 2000, units = "px")

```

## Poisson Regression
### Model Fitting
```{r}
## Poisson Regression
### Simple
poisson_simple_resampled_illness <- poisson_workflow_simple_illness %>%
  fit_resamples(resamples = illness_fold,
    control = control_resamples(save_pred = TRUE))

### Multivariate
poisson_complex_resampled_illness <- poisson_workflow_complex_illness %>%
  fit_resamples(resamples = illness_fold,
    control = control_resamples(save_pred = TRUE))
```

### Model Evaluation
```{r}
# Metrics
## Simple
poisson_simple_resampled_illness %>%
  collect_metrics()
## Complex
poisson_complex_resampled_illness %>%
  collect_metrics()

# Check Parameters
## Simple
poisson_simple_fit_illness <- poisson_workflow_simple_illness %>%
  fit(illness_data) %>%
  tidy() %>%
  arrange(-estimate)
## Complex
poisson_complex_fit_illness <- poisson_workflow_complex_illness %>%
  fit(illness_data) %>%
  tidy() %>%
  arrange(-estimate)

# Check Variable Importance
## Simple
poisson_workflow_simple_illness %>%
  fit(illness_data) %>%
  extract_fit_engine() %>%
  vip()
## Complex
poisson_workflow_complex_illness %>%
  fit(illness_data) %>%
  extract_fit_engine() %>%
  vip()

# Plotting Predictions vs Illnesses
## Simple
poisson_simple_resampled_illness %>%
  collect_predictions() %>%
  # filter(Illnesses < 300) %>%
  ggplot(aes(Illnesses, .pred, color = id2, group = id2)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  geom_abline(intercept = 0, slope = 1) +
  labs(y = "Predicted Illnesses", title = "Predicted Illnesses VS Illnesses by Simple Poisson Regression")

ggsave("Predicted Illnesses VS Illnesses by poisson regression with main predictor.png", path = images_stats_path, width = 3000, height = 2000, units = "px")

## Complex
poisson_complex_resampled_illness %>%
  collect_predictions() %>%
  # filter(Illnesses < 500) %>%
  ggplot(aes(Illnesses, .pred, color = id2, group = id2)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  geom_abline(intercept = 0,slope = 1) +
  labs(y = "Predicted Illnesses", title = "Predicted Illnesses VS Illnesses by linear regression with all predictors")

ggsave("Predicted Illnesses VS Illnesses by poisson regression with all predictors.png", path = images_stats_path, width = 3000, height = 2000, units = "px")

```
According to the result, the model built using multivariate linear regression seems to have the smallest root mean squared error and highest R^2. However, none of those models can catch outliers (Illness > 1000).

# Modeling Hospitilizations

To create a workflow for the Poisson regression model, we will use the default *glm* engine again.

```{r}
# Cross-Validation
hosp_fold <- vfold_cv(hosp_data, repeats = 2, strata = Hospitalizations)

# Data Preprocessing
## Simple Regression Recipe
hosp_simple_recipe <- recipe(Hospitalizations ~ Simplified_Etiology, data = hosp_data) %>%
  step_dummy(Simplified_Etiology)

## Multivariate Regression Recipe
hosp_complex_recipe <- recipe(Hospitalizations ~ ., data = hosp_data) %>%
  preprocessing_recipe()

## Recipe for all predictors
### Lump small proportion into "other" using step_other
### Create dummy code for nominal variables
# hosp_all_rec=recipe(Hospitalizations ~.,data=hosp_data)%>%
#  step_other(State,Location,`IFSAC Category`, Simplified_Etiology,threshold = 0.01)%>%
#  step_dummy(all_nominal_predictors())

# Check Preprocessed Dataset
hosp_complex_recipe %>%
  prep() %>%
  bake(new_data = NULL)

# Create Workflows
## Simple
linear_workflow_simple_hosp <- workflow() %>%
  add_recipe(hosp_simple_recipe) %>%
  add_model(linear_model)
poisson_workflow_simple_hosp <- workflow() %>%
  add_recipe(hosp_simple_recipe) %>%
  add_model(poisson_reg())
## Multivariate
linear_workflow_complex_hosp <- workflow() %>%
  add_recipe(hosp_complex_recipe) %>%
  add_model(linear_model)
poisson_workflow_complex_hosp <- workflow() %>%
  add_recipe(hosp_complex_recipe) %>%
  add_model(poisson_reg())
```

## Linear Regression
```{r}
set.seed(234)
```

### Model Fitting
```{r}
# Simple
lm_simple_resampled_hosp <- linear_workflow_simple_hosp %>%
  fit_resamples(resamples = hosp_fold,
    control = control_resamples(save_pred = TRUE))

# Complex
lm_complex_resampled_hosp <- linear_workflow_complex_hosp %>%
  fit_resamples(resamples = hosp_fold,
    control = control_resamples(save_pred = TRUE))
```

### Model Evaluation
```{r}
# Metrics
lm_simple_resampled_hosp %>%
  collect_metrics()
lm_complex_resampled_hosp %>%
  collect_metrics()

# Check Parameters
linear_workflow_simple_hosp %>%
  fit(hosp_data) %>%
  tidy() %>%
  arrange(-estimate)
linear_workflow_complex_hosp %>%
  fit(hosp_data) %>%
  tidy() %>%
  arrange(-estimate)

## Visualizing Variable Importance
linear_workflow_simple_hosp %>%
  fit(hosp_data) %>%
  extract_fit_engine() %>%
  vip()
linear_workflow_complex_hosp %>%
  fit(hosp_data) %>%
  extract_fit_engine() %>%
  vip()

# Plotting predictions vs Hospitalizations
lm_simple_resampled_hosp %>%
  collect_predictions() %>%
  # filter(Hospitalizations < 50) %>%
  ggplot(aes(Hospitalizations, .pred, color = id2))+
  geom_point() +
  geom_smooth(se = FALSE) +
  geom_abline(intercept = 0, slope = 1) +
  labs(y = "Predicted Hospitalizations", 
       title = "Predicted Hospitalizations VS Hospitalizations \by Simple Linear Regression", color = "Fold")

ggsave("Predicted Hospitalizations VS Hospitalizations by linear regression with main predictor.png", path = images_stats_path, width = 3000, height = 2000, units = "px")

lm_complex_resampled_hosp %>%
  collect_predictions() %>%
  #filter(Hospitalizations < 50) %>%
  ggplot(aes(Hospitalizations, .pred, color = id2))+
  geom_point() +
  geom_smooth(se = FALSE) +
  geom_abline(intercept = 0, slope = 1) +
  labs(y = "Predicted Hospitalizations", 
       title="Predicted Hospitalizations VS Hospitalizations \by Multivariate Linear Regression", color = "Fold")

ggsave("Predicted Hospitalizations VS Hospitalizations by linear regression with all predictors.png", path = images_stats_path, width = 3000, height = 2000, units = "px")
```
## Poisson Regression
### Model Fitting
```{r}
## Poisson Regression
### Simple
poisson_simple_resampled_hosp <- poisson_workflow_simple_hosp %>%
  fit_resamples(resamples = hosp_fold,
    control = control_resamples(save_pred = TRUE))

### Multivariate
poisson_complex_resampled_hosp <- poisson_workflow_complex_hosp %>%
  fit_resamples(resamples = hosp_fold,
    control = control_resamples(save_pred = TRUE))
```

### Model Evaluation
```{r}
# Metrics
poisson_simple_resampled_hosp %>%
  collect_metrics()
poisson_complex_resampled_hosp %>%
  collect_metrics()

# Check Parameters
poisson_workflow_simple_hosp %>%
  fit(hosp_data) %>%
  tidy() %>%
  arrange(-estimate)

poisson_workflow_complex_hosp %>%
  fit(hosp_data) %>%
  tidy() %>%
  arrange(-estimate)

# Variable Importance
poisson_workflow_simple_hosp %>%
  fit(hosp_data) %>%
  extract_fit_engine() %>%
  vip()
poisson_workflow_complex_hosp %>%
  fit(hosp_data) %>%
  extract_fit_engine() %>%
  vip()

# Predictions vs Hospitalizations
poisson_simple_resampled_hosp %>%
  collect_predictions() %>%
  # filter(Hospitalizations < 50) %>%
  ggplot(aes(Hospitalizations, .pred, color = id2)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  geom_abline(intercept = 0,slope = 1) +
  labs(y = "Predicted Hospitalizations", 
       title = "Predicted Hospitalizations VS Hospitalizations \by Simple Linear Regression", color = "Fold")

ggsave("Predicted Hospitalizations VS Hospitalizations by poisson regression with main predictor.png", path = images_stats_path, width = 3000, height = 2000, units = "px")

poisson_complex_resampled_hosp %>%
  collect_predictions() %>%
  # filter(Hospitalizations < 50) %>%
  ggplot(aes(Hospitalizations, .pred, color = id2)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  geom_abline(intercept = 0, slope = 1) +
  labs(y = "Predicted Hospitalizations", title = "Predicted Hospitalizations VS Hospitalizations \by Multivariate Linear Regression", color = "Fold")

ggsave("Predicted Hospitalizations VS Hospitalizations by poisson regression with all predictors.png", path = images_stats_path, width = 3000, height = 2000, units = "px")

```
According to the results, the model built using multivariate linear regression seems to have the smallest root mean squared error and highest R^2 (14%). But again, outliers (Hospitalizations > 100) cannot be predicted well by any of those models.

# Machine Learning Approach
```{r}
set.seed(345)
```

## Random Forest
### Predicting First Outcome `Illnesses` Using All Predictors
```{r}
# Model Specification
rf_spec <- rand_forest(mtry = tune(), min_n = tune(), trees = 1000) %>%
  set_engine("ranger", importance = 'impurity') %>%
  set_mode("regression")

# illness_all_rec <- recipe(Illnesses~.,data=illness_data) %>%
#  step_other(State,Location,`IFSAC Category`,Simplified_Etiology,threshold = 0.01)%>%
#  step_dummy(all_nominal())

# Workflow
random_forest_illness_workflow <- workflow() %>%
  add_recipe(illness_complex_recipe) %>%
  add_model(rf_spec)

# Training and Tuning Model
doParallel::registerDoParallel()
random_forest_illness_tuned <- random_forest_illness_workflow %>%
  tune_grid(
    resamples = illness_fold,
    grid = 15,
    control = control_grid(save_pred = TRUE),
    metrics = metric_set(rmse)
  )
```

### Model Evaluation
```{r}
# Plot
random_forest_illness_tuned %>%
  autoplot()

# Final Specifications
random_forest_illness_tuned_best <- random_forest_illness_tuned %>%
  select_best()
random_forest_illness_final <- random_forest_illness_workflow %>%
  finalize_workflow(random_forest_illness_tuned_best)
random_forest_illness_fit <- random_forest_illness_final %>%
  fit(illness_data)

# Plotting Predicted Illness VS Illnesses
random_forest_illness_predict <- random_forest_illness_fit %>%
  predict(illness_data)

illness_data %>%
  select(Illnesses) %>%
  bind_cols(random_forest_illness_predict) %>%
  ggplot(aes(Illnesses, .pred)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1)+
  geom_smooth(se = FALSE) +
  labs(y = "Predicted Illnesses", 
       title = "Predicted Illnesses VS Illnesses by Random Forest Model")
ggsave("Predicted illnesses VS illnesses by random forest model.png", path = images_stats_path, width = 3000, height = 2000, units = "px")
```

### Predicting Second Outcome `Hospitalizations` Using All Predictors
```{r}
set.seed(456)

# Creating workflow
random_forest_hosp_workflow <- workflow() %>%
  add_recipe(hosp_complex_recipe) %>%
  add_model(rf_spec)
# Training and tuning model
doParallel::registerDoParallel()
random_forest_hosp_tuned <- random_forest_hosp_workflow %>%
  tune_grid(
    resamples = hosp_fold,
    grid = 15,
    control = control_grid(save_pred = TRUE),
    metrics = metric_set(rmse)
  )
```

### Model Evaluation
```{r}
# Plot
random_forest_hosp_tuned %>%
  autoplot()

# Final Specifications
random_forest_hosp_tuned %>%
  show_best()
random_forest_hosp_best <- random_forest_hosp_tuned %>%
  select_best()

random_forest_hosp_final <- random_forest_hosp_workflow %>%
  finalize_workflow(random_forest_hosp_best)

random_forest_hosp_fit <- random_forest_hosp_final %>%
  fit(hosp_data)

random_forest_hosp_predict <- random_forest_hosp_fit %>%
  predict(hosp_data)

hosp_data %>%
  select(Hospitalizations) %>%
  bind_cols(random_forest_hosp_predict) %>%
  ggplot(aes(Hospitalizations, .pred)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1) +
  geom_smooth(se = FALSE) +
  labs(y = "Predicted Hospitalizations",
       title = "Predicted Hospitalizations VS Hospitalizations by Random Forest")

ggsave("Predicted Hospitalizations VS Hospitalizations by random forest model.png", path = images_stats_path, width = 3000, height = 2000, units = "px")
```

Overall, random forest model has the lowest rmse among those selected models when training data is fitted; Also, the random forest model outperforms other models in predicting outliers.

## Bagging
### Illnesses
```{r}
bag_training_illness <- bagging(formula = Illnesses ~ .,
                        data = illness_data,
                        method = "treebag",
                        trControl = trainControl(method = "cv", number = 30),
                        nbagg = 500,
                        coob = TRUE,
                        control = rpart.control(minsplit = 2, cp = 0))

bag_training_illness
```

### Hospitalizations
```{r}
bag_training_hosp <- bagging(formula = Hospitalizations ~ .,
                        data = hosp_data,
                        method = "treebag",
                        trControl = trainControl(method = "cv", number = 30),
                        nbagg = 500,
                        coob = TRUE,
                        control = rpart.control(minsplit = 2, cp = 0))

bag_training_hosp
```

## Organize Model Performance and Choose the Best One
### Illnesses
```{r}
# Linear Regression
lm_complex_resampled_illness %>%
  collect_metrics()

# Poisson Regression
poisson_complex_resampled_illness %>%
  collect_metrics()

# Random Forest
random_forest_illness_predict %>%
  collect_metrics()

# Make Dataframe of Model RMSEs

```

### Hospitalizations
```{r}
# Linear Regression
# Poisson Regression
# Random Forest

```

## Fit Testing Data