### 2023-04-21 code/analysis_code/exploratory_analysis.qmd
more work

---
title: "Exploratory analysis script"
author: "Weifan Wu and Kailin Chen"
date: "10/28/2021"
output: html_document
---

# Load Libraries
```{r}
# General Data Handling
library(tidyverse)

# For Data Loading and Saving
library(here)
```

# Load Data

## Training Data
```{r}
# Path to Data
data_location <- here(
  "data","processed_data","processed_training_data.rds")
# Import Data
training_data_final <- readRDS(data_location)
```

## Testing/Analysis Data
```{r}
# Path to Data
data_location <- here(
  "data","processed_data","processed_testing_data.rds")
# Import Data
analysis_data_final <- readRDS(data_location)
```

# Data Exploration
## Training Data

The outcome of interest we will explore for the training data is hospitalizations.
### Template Setup
```{r}
# Template Graph
theme_set(theme_classic())
theme_update(plot.title = element_text(face = "bold", size = 12))
theme_update(plot.title = element_text(hjust = 0.5))
```

### Pathogen Type + Etiology vs Hospitalization
```{r}
# Pathogen Type vs Hospitalization
ggplot(training_data_final %>% 
         group_by(Pathogen_Type, Hospitalizations) %>% 
         count() %>% mutate(true_hospitalization = n * Hospitalizations), 
       aes(x = Pathogen_Type, y = true_hospitalization, 
           )) +
  geom_col() + 
  labs(x = "Pathogen Type", y = "Hospitalizations", 
       title = "Bar Chart of Hospitalizations by Pathogen type") +
  theme_classic() + 
  theme(plot.title = element_text(face = "bold", size = 11, hjust = 0.5))

## Save Graph
image_path <- here("results","images_EDA")
ggsave("pathogen_type_v_hospitalizations.png", path = image_path, width = 3000, height = 2000, units = "px")
```

### Time vs Hospitalizations
```{r}
# Data Visualization
## Illnesses
# ggplot(training_data_final %>%
#          group_by(Year, Pathogen_Type, Illnesses) %>%
#          count() %>%
#          mutate(true_illnesses = n * Illnesses),
#        aes(x = Year, y = true_illnesses,
#            group = Pathogen_Type, color = Pathogen_Type)) +
#   geom_point() + geom_jitter() +
#   labs(x = "Year", y = "Illnesses",
#        title = "Scatter Plot of Illnesses by Year and Pathogen Type",
#        color = "Pathogen Type") +
#   facet_wrap(~ Pathogen_Type) +
#   theme(legend.position = "none")

## Hospitalizations VS year
ggplot(training_data_final %>% 
         group_by(Year, Pathogen_Type, Hospitalizations) %>% 
         count() %>% 
         mutate(true_hospitalization = n * Hospitalizations), 
       aes(x = Year, y = true_hospitalization, 
           group = Pathogen_Type, color = Pathogen_Type)) + 
  geom_point() + geom_jitter() + 
  labs(x = "Year", y = "Hospitalizations", 
       title = "Scatter Plot of Hospitalizations by Year and Pathogen Type",
       color = "Pathogen Type") + 
  facet_wrap(~ Pathogen_Type) + 
  theme(legend.position = "none")

## Save Graph
ggsave("year_v_hospitalization.png", path = image_path, width = 3000, height = 2000, units = "px")

## Hospitalizations VS month
ggplot(training_data_final %>% 
         group_by(Numeric_Month, Hospitalizations) %>% 
         count() %>% 
         mutate(true_hospitalization = n * Hospitalizations), 
       aes(x = as.factor(Numeric_Month), y = true_hospitalization)) +
  geom_boxplot() +
  labs(x = "Month", y = "Hospitalizations", 
       title = "Scatter Plot of Hospitalizations by Month") + 
  theme(legend.position = "none")

## Save Graph
ggsave("month_v_hospitalization.png", path = image_path, width = 3000, height = 2000, units = "px")
```

### Bacterial Etiologies for Outbreaks
```{r}
# General
ggplot(training_data_final %>% 
         select(Pathogen_Type, Simplified_Etiology, Hospitalizations) %>%
         filter(Pathogen_Type == "Bacteria")%>%
         group_by(Simplified_Etiology)%>%
         mutate(all_hospitalizations=sum(Hospitalizations)),
       aes(x = fct_reorder(Simplified_Etiology,all_hospitalizations))) + geom_bar() + 
  labs(x = "Bacteria", y = "Hospitalizations", 
       title = "Bar Chart of Bacteria Frequency") + coord_flip()

ggsave("bacterial_etiology_v_hospitalization_frequency.png", path = image_path, width = 3000, height = 2000, units = "px")

#  Top 3 Bacterial Etiology: Salmonella, E. coli, and Clostridium
Top_Hosp <- training_data_final %>%
  select(Simplified_Etiology, Hospitalizations, `IFSAC Category`) %>%
  group_by(Simplified_Etiology, Hospitalizations, `IFSAC Category`) %>%
  filter(Simplified_Etiology %in% c("Salmonella","Clostridium","E. coli")) %>%
  filter(`IFSAC Category` != "Multiple") %>%
  filter(`IFSAC Category` != "Other") %>%
  count() %>%
  mutate(total_hospitalizations = n * Hospitalizations)

ggplot(Top_Hosp, aes(x=fct_reorder(`IFSAC Category`, total_hospitalizations),y = total_hospitalizations))+
  geom_boxplot() + labs(x = "Food Type", y = "Total Hospitalizations", 
                        title = "Hospitalizations by Food Type for Top 3 Bacterial Etiology") +
  facet_wrap(vars(Simplified_Etiology))+
  coord_flip()
ggsave("top3_bacterial_etiology_v_food_hospitalization_frequency.png", path = image_path, width = 3000, height = 2000, units = "px")
```

### Hospitalizations by State
```{r}
# States vs Hospitalization
state_v_hospitalization_training <- training_data_final %>% 
  group_by(State) %>% 
  mutate(all_hospitalizations = sum(Hospitalizations)) %>% 
  count(all_hospitalizations) %>% 
  select(State, all_hospitalizations) %>% arrange(desc(all_hospitalizations))

training_data_final %>% 
  group_by(State) %>% 
  mutate(all_illnesses = sum(Illnesses)) %>% 
  count(all_illnesses) %>% 
  select(State, all_illnesses) %>% arrange(desc(all_illnesses))

## Save Table
table_path <- here("results", "state_v_hospitalization_training.rds")
saveRDS(state_v_hospitalization_training, file = table_path)

rmarkdown::paged_table(state_v_hospitalization_training)

```

### State Map of Hospitalizations
```{r}
training_hosp <- training_data_final%>%
  mutate(state = tolower(State))%>%
  group_by(state)%>%
  mutate(all_hosp = sum(Hospitalizations)) %>% count(all_hosp) %>%
  select(state, all_hosp)
training_hosp

state_map <- training_hosp %>%
  distinct(state)%>%
  inner_join(., map_data("state"), by = c("state"="region"))

state_map

map_final <- right_join(training_hosp,state_map,by=c("state"="state"))

map_final %>%
  ggplot(aes(x=long, y=lat, 
             fill = all_hosp, group = group)) +
  geom_polygon(color="black", size = 0.1) +
  geom_polygon(data = state_map, aes(x= long, y= lat, group = group), 
               fill = NA, color = "black", size= 0.3) +
  scale_fill_gradient(name = "Number of \n Hospitalizations",
                      low = "#FFFFFF", high="#FF0000",
                       limits = c(0,NA)) +
  labs(title = "Cases of Hospitalizations Caused by Foodborne Illnesses in Each State (1998-2015)") +
  theme_void() + theme(plot.title = element_text(hjust = 0.5)) + 
  theme(plot.title = element_text(face = "bold", size = 12)) +
  coord_quickmap()

ggsave("State_hospitalization_map.png", path = image_path, width = 3000, height = 2000, units="px")
```

### Food Type and Hospitalizations
```{r}
# Food Type and Hospitalizations
ggplot(training_data_final %>% 
         filter(`IFSAC Category` != "Multiple") %>% 
         filter(`IFSAC Category` != "Other") %>%
         group_by(`IFSAC Category`) %>% 
         mutate(all_hospitalizations = sum(Hospitalizations)) %>%
         count(all_hospitalizations), 
       aes(x = fct_reorder(`IFSAC Category`,all_hospitalizations), y = all_hospitalizations, 
           )) + 
  geom_col() + 
  labs(x = "Food Type", y = "Hospitalizations", title = "Hospitalizations by Food Type, Excluding 'Multiple' and 'Other'") +
  coord_flip() + theme(legend.position = "none")

ggsave("Food_type_v_hospitalization.png", path = image_path, width = 3000, height = 2000, units = "px")

# Extra: Food Type & Illnesses
training_data_food_and_illnesses <- training_data_final %>% 
         filter(`IFSAC Category` != "Multiple") %>% 
         filter(`IFSAC Category` != "Other") %>%
         group_by(`IFSAC Category`) %>% 
         mutate(all_illnesses = sum(Illnesses)) %>%
         count(all_illnesses) %>% 
  arrange(desc(all_illnesses))
table_path <- here("results", "training_food_vs_illnesses.rds")
saveRDS(training_data_food_and_illnesses, file = table_path)

```

### Locations and Hospitalizations
```{r}
Location_v_Hospitalization <- training_data_final%>%
  group_by(Location)%>%
  mutate(all_hospitalizations = sum(Hospitalizations)) %>%
         count(all_hospitalizations)%>%
  ggplot(aes(x = fct_reorder(Location,all_hospitalizations), y = all_hospitalizations)) +
  geom_col() + 
  labs(x = "Locations", y = "Hospitalizations", title = "Outbreak Location versus Hospitalizations") +
  coord_flip()

Location_v_Hospitalization

ggsave("Location_v_hospitalization.png", path = image_path, width = 3000, height = 2000, units = "px")

```

All outbreaks with a high hospitalization count (> 200 cases) are caused by bacteria. It also appears that most outbreaks caused by bacteria are due to Salmonella. California and Minnesota are the two states with the highest total hospitalization count due to foodborne illness from 1998 to 2015. Restaurants and Private home are the most frequent locations with the highest total hospitalization count.

## Testing Data

### Pathogen Type vs Illnesses Over Time
```{r}
ggplot(analysis_data_final %>% 
         group_by(Year, Pathogen_Type, Illnesses) %>% 
         count() %>% 
         mutate(true_illness_count = n * Illnesses), 
       aes(x = Year, y = true_illness_count, group = Pathogen_Type, color = Pathogen_Type)) + 
  geom_point() + 
  geom_jitter() + 
  labs(x = "Year", y = "Cases", title = "Scatter Plot of Illness by Year and Pathogen Type")

ggsave("time_v_illness_pathogen_type.png", path = image_path, width = 3000, height = 2000, units = "px")

```

### Pathogen Type and Etiology vs Illnesses
```{r}
# Visualization
ggplot(analysis_data_final %>% 
         group_by(Pathogen_Type, Simplified_Etiology, Illnesses) %>% 
         count() %>% 
         mutate(true_illness_count = n * Illnesses), 
       aes(x = Pathogen_Type, y = Illnesses, group = Simplified_Etiology, fill = Simplified_Etiology)) + 
  geom_col() + 
  labs(x = "Pathogen Type", y = "Cases", title = "Stacked Bar Chart of Illnesses by Pathogen Type and Etiology")

# Save Graph
ggsave("pathogen_type_etiology_v_illness.png", path = image_path, width = 3000, height = 2000, units = "px")
```

### Bacterial Etiologies
```{r}
ggplot(analysis_data_final %>% 
         group_by(Simplified_Etiology) %>% 
         mutate(Illnesses=sum(Illnesses))%>%
         filter(Pathogen_Type == "Bacteria"), aes(x = fct_reorder(Simplified_Etiology,Illnesses))) + 
  geom_bar() + 
  labs(x = "Bacteria", y = "Cases", title = "Bar Chart of Bacteria Frequency for Illnesses")+
  coord_flip()

# Save Graph
ggsave("bacterial_etiology_v_outbreak_frequency.png", path = image_path, width = 3000, height = 2000, units = "px")
```

### Illnesses by State
```{r}
state_v_illness_analysis <- analysis_data_final %>% 
  group_by(State) %>%
  mutate(all_illnesses = sum(Illnesses)) %>%
  count(all_illnesses) %>% 
  select(State, all_illnesses) %>% 
  arrange(desc(all_illnesses))

# Save Table
table_path <- here("results", "state_v_illness_training.rds")
saveRDS(state_v_illness_analysis, file = table_path)

rmarkdown::paged_table(state_v_illness_analysis)
```

### Food Type and Illnesses
```{r}
analysis_data_final %>% filter(!`IFSAC Category` %in% c("Multiple", "Other"))%>%
  filter(Illnesses>0)%>%
  group_by(`IFSAC Category`)%>%
  mutate(Illnesses=sum(Illnesses))%>%
ggplot(aes(x = fct_reorder(`IFSAC Category`,Illnesses), y = Illnesses)) + 
  geom_col() + 
  scale_y_log10()+
  labs(x = "Food Type", y = "Illnesses", title = "Illnesses by Food Type, Excluding 'Multiple' and 'Other'")+
  coord_flip()

ggsave("food_type_v_illness_frequency.png", path = image_path, width = 3000, height = 2000, units = "px")

```

### Locations and Illnesses
```{r}
# Locations and Illnesses
ggplot(analysis_data_final %>% group_by(Location) %>% 
         mutate(all_illnesses = sum(Illnesses)) %>%
         count(all_illnesses), aes(x = fct_reorder(Location,all_illnesses), y = all_illnesses)) +
  geom_col() + 
  labs(x = "Locations", y = "Illnesses", title = "Illnesses by Location")+
  coord_flip()

# Save Graph
ggsave("Location_v_Illness.png", path = image_path, width = 3000, height = 2000, units = "px")
```
The outbreaks with a high number of cases (> 500) are caused by bacteria. The stacked barplot indicates that the top two etiologic agents that contribute to fooborne-illnesses are Salmonella and Norovirus, which belong to the pathogen type of Bacteria and Virus respectively. In bacterial pathogen, Salmonella genus is the most dominant factor that is associated with foodborne outbreaks, followed by E. coli species and Vibrio genus. Geographically speaking, Arkansas and Minnesota are the two states with the highest accumulated foodborne illnesses from 1998 to 2015. Salad, Poultry, and Fruits are the three most common types of food vectors that are involved in transmitting foodborne pathogens. Resturants and Private home are the most frequent locations that foodborne outbreaks occur.

Overall, the relationships between predictors and outcomes of training dataset share similar patterns with those of analysis dataset.