### 2023-04-28 reviews/Project_Review_DeannaL.qmd
combine two reviews into review folder

---
title: Project Review Weifan and Kai
author: Deanna Lanier
date: "`r file.mtime(knitr::current_input())`"
format: 
  html:
    toc: true
    toc-depth: 3
    number-sections: true
---

# Overview

Title of project: Foodborne-Outbreak

Name of project author(s): Weifan Wu & Kailin (Kai) Chen

Name of project reviewer: Deanna Lanier

# Specific project content evaluation
Evaluate the different parts of the project by filling in the sections below.

## Background, Context and Motivation
How well is the context of the project described? Is a comprehensive background, including summary of previous/related work given? Is the project well placed into the context of existing work (including proper referencing of existing work). Is it clear why the project was undertaken and what new information it hopes to provide?

### Feedback and Comments

The background does a good job of describing your topic and the previous related studies. There a relevant references and it is well documented why the project was undertaken and the new information it will provide

### Summary assessment (PICK ONE, DELETE THE OTHERS)
* strong contextualization and motivation

## Question description
How well and clear are the question(s)/hypotheses the project aims to address described? Is it clear how the questions relate to the data?

### Feedback and Comments

The questions and hypothesis were clearly stated and well explained. They relate to the data and the background well and are applicable to the analysis.

### Summary assessment
* question/hypotheses fully clear

## Data description
How well is the data overall described? Is the source provided? Is a codebook or other meta-information available that makes it clear what the data is?

### Feedback and Comments

The data are described in the data sources section and clearly explain where the data were obtained. There is no meta-information provided on the data sources so it is hard to be clear on what the variables are. I would recommend including an additional spreadsheet with your data dictionary so it is easier for other users to interpret.

### Summary assessment
* source and overall structure of data somewhat explained

## Data wrangling and exploratory analysis
How well is the data cleaned/processed and explored? Are all steps reasonable and well explained? Are alternatives discussed and considered? Are meaningful exploratory results shown (e.g. in the supplementary materials)?

### Feedback and Comments

The data processing was very comprehensive and well explained. I especially liked the Pareto Chart with Missing Values figures in your processing script. I think all of the steps were reasonable. For the exploratory analysis, the figures look great and all the results presented are related to the overarching study. The meaningful results from the exploration are explained and documented.

### Summary assessment
* essentially no weaknesses in wrangling and exploratory component

## Appropriateness of Analysis
Were the analysis methods appropriate for the data? Was the analysis done properly? Were different components of the analysis (e.g. performance measure, variable selection, data pre-processing, model evaluation) done in the best way possible and explained well?

### Feedback and Comments

Great analysis! You started with explain the findings from our exploratory analysis and linked that to the rational behind your statistical analysis which was very helpful. You this statistical analysis was very appropriate for the questions and hypothesis and you utilized multiple models and included an great model interpretation with some performance metrics. The model results were also explained well and additional preprocessing steps were described.

### Summary assessment
* strong and reasonable analysis

## Presentation
How well are results presented? Are tables and figures easy to read and understand? Are the main figures/tables publication level quality?

### Feedback and Comments

The figures are well presented. They include legible titles, descriptions, and are relevant to the analysis. The axis titles could be slightly larger but overall, they are great quality. I would recommend making the titles of your figure have consistent formatting. Some include the figure number and others do not and the last figure did not include a title at all. For the tables, I would recommend having all of the captions and formatting done in a different .qmd or .r file so that we do not see the code you used to add the caption and section criteria.

### Summary assessment
* results are presented ok, with room for improvement

## Discussion/Conclusions
Are the study findings properly discussed? Are strengths and limitations acknowledged? Are findings interpreted properly?

### Feedback and Comments

The findings are proper discussed. The hypothesis stated earlier in the manuscript were mentioned in the discussion as to whether the analysis supported or did not support them. I also liked that other findings from previous studies were included and related to the results from this project. There was a dedicated section for strengths and limitations, however it seemed to focus more on future directions of the study. You should definitely keep the future directions, but add the strengths and limitations. The conclusion was concise and all of the findings seem to be interpreted properly.

### Summary assessment
* strong, complete and clear discussion

## Further comments

Overall I think the project is great and I love the topic and dataset you all chose. This is a great analysis and it seems like a lot of consideration was put into the methods of analysis and the model interpretation. Great work!

# Overall project content evaluation
Evaluate overall features of the project  by filling in the sections below.

## Structure
Is the project well structured? Are files in well labeled folders? Do files have reasonable names? Are all "junk" files not needed for analysis/reproduction removed? By just looking at files and folders, can you get an idea of how things fit together?

### Feedback and Comments

The project repository is well structured and all of the files and folders are given reasonable names. There are no junk files or folders that do not need to be there. Even without the instructions in the readme, it was very easy to know how everything fit together and which code to run first just based on the names.

### Summary assessment
* well structured

## Documentation 
How well is the project documented? Are you able to understand each step of the whole analysis, each decision that was made, and each line of code? Is enough information provided as comments in code or as part of Rmd files?

### Feedback and Comments

This project could benefit from a lot more comments in the scripts. There are very few comments that explain what you are doing in the code chunks and that would be something helpful to add.

### Summary assessment
* decently documented with some gaps

## Reproducibility
Are all results fully reproducible? Is documentation provided which clearly explains how to reproduce things, and does it work without the need for any manual intervention? Are you able to re-run the whole analysis without having to do manual interventions/edits?

### Feedback and Comments

The README.md files could be more comprehensive. The one in the primary folder gives no information on the steps required to reproduce the results and the ones in the sub folders are not helpful either. All of the results are fully reproducible and there were no manual additions required for the scripts to run in their entirety. The only error I had was with the last lines in the statistical analysis (starting at line 756) which was saving the files. This did not affect any of the manuscript or figures though. I would fix that chunk so there are no errors and all results save properly.

### Summary assessment
* fully reproducible without issues

## Thoroughness
How thorough was the overall study? Were alternatives (e.g. different ways of processing the data or different models) considered? Were alternatives discussed? Were the questions/hypotheses fully and thoroughly addressed?

### Feedback and Comments

This was a thorough study overall. Four different models were compared for the analysis and they used the performance metrics to determine the best model for analysis. The questions and hypotheses were all thoroughly addressed in the discussion and conclusion section. They found that 2 of the 3 hypotheses were not supported after the analysis and all of the questions were addressed through out the project.

### Summary assessment
* strong level of thorougness

## Further comments

Overall this is a great analysis and very comprehensive. My only suggestions would be to adjust your axis labels and title sizes and to incorporate more descriptive README.md files to make reproducing your results more easier. Great work!